---
title: "Prediction of Patient Status using Voice Measurements"
author: "INSOFE Lab Assignment on Logistic Regression - Batch 30"
date: "8 July 2017"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
---

**NOTE** Before starting this assignment please remember to clear your environment, you can do that by running the following code chunk

```{r}
rm(list = ls(all=TRUE))


```

# Agenda 

* Get the data

* Data Pre-processing

* Build a model

* Predictions

* Communication


# Reading & Understanding the Data

* The "parkinsons_data.csv"" dataset is composed of a range of biomedical voice measurements. Each column in the table is a particular voice measure, and each row corresponds one of 195 voice recordings.

* The dataset has 195 rows and 23 columns. 

* The column/variable names' explanation is given below:

1) __MDVP:Fo(Hz) :__ Average vocal fundamental frequency

2) __MDVP:Fhi(Hz) :__ Maximum vocal fundamental frequency

3) __MDVP:Flo(Hz) :__ Minimum vocal fundamental frequency

4) __MDVP:Jitter(%); MDVP:Jitter(Abs); MDVP:RAP; MDVP:PPQ; Jitter:DDP :___ Several measures of variation in fundamental frequency

5) __MDVP:Shimmer; MDVP:Shimmer(dB); Shimmer:APQ3; Shimmer:APQ5; MDVP:APQ; Shimmer:DDA :__ Several measures of variation in amplitude

6) __NHR; HNR :__ Two measures of ratio of noise to tonal components in the voice

7) __status :__ Health status of the subject (Parkinson's) - Positive, (Normal) - Healthy

8) __RPDE; D2 :__  Two nonlinear dynamical complexity measures

9) __DFA :__ Signal fractal scaling exponent

10) __spread1; spread2; PPE :__ Three nonlinear measures of fundamental frequency variation

* Make sure the dataset is located in your current working directory and read in the data

```{r}
setwd("C:\\insofe\\Logistic Regression")
getwd()

parkinsons_data <- read.csv("parkinsons_data.csv", header=T)

```

* Use the str() function to get a feel for the dataset.

```{r}
str(parkinsons_data)

```


* Take a look at the data using the "head()" and "tail()" functions

```{r}

head(parkinsons_data)
tail(parkinsons_data)

```

* Are there any missing values in the dataset?

```{r}
sum(is.na(parkinsons_data))

```

# Data Pre-processing

## Train/Test Split

* Split the data 70/30 into train and test sets, using __Stratified Sampling__ by setting the seed as "786"

```{r}


set.seed(786)
library(caret)
parkinsons_rows <- createDataPartition(parkinsons_data$status, p = 0.7, list = F)

trainP_data <- parkinsons_data[parkinsons_rows, ]

testP_data <- parkinsons_data[-parkinsons_rows, ]

summary(trainP_data)
summary(testP_data)

str(testP_data)

```

# Build a model

## Basic Logistic Regression Model

* Use the glm() function to build a basic model

* Build a model using all the variables, excluding the response variable, in the dataset

```{r}


log_parkinsons = glm(status~., data =trainP_data, family = binomial)

```

* Get the summary of the model and understand the output

```{r}
summary(log_parkinsons)

```
* Improve the model using stepAIC.
```{r}
library(MASS)
log_parkinsons_step = stepAIC(log_parkinsons)

#stepAIC improves the AIC value
```

* Use vif to find any multi-collinearity
```{r}
library(car)
log_parkinsons_step_vif = vif(log_parkinsons_step)
log_parkinsons_step_vif

```

* Check the model summary to check for any insignificant variables
```{r}
summary(log_parkinsons_step)


#Since AIC value is less, we choose 'log_relog_parkinsons_step' as our final model
```

## Create an ROC Curve

1) Get a list of predictions (probability scores) using the predict() function

```{r}

prob_train <- predict(log_parkinsons_step, type = "response")

```


2) Using the ROCR package create a "prediction()" object

```{r}
library(ROCR)

pred <- prediction(prob_train, trainP_data$status)

```


3) Extract performance measures (True Positive Rate and False Positive Rate) using the "performance()" function from the ROCR package

```{r}

perf <- performance(pred, measure="tpr", x.measure="fpr")

```


4) Plot the ROC curve using the extracted performance measures (TPR and FPR)

```{r}

plot(perf, col=rainbow(10), colorize=T, print.cutoffs.at=seq(0,1,0.1))

```

* Extract the AUC score of the ROC curve and store it in a variable named "auc"

```{r}

perf_auc <- performance(pred, measure="auc")

# Access the auc score from the performance object

auc <- perf_auc@y.values[[1]]

print(auc)


```


## Choose a Cutoff Value

* Based on the trade off between TPR and FPR depending on the business domain, a call on the cutoff has to be made.

```{r}

### Write your answer here
pred_class <- ifelse(prob_train> 0.3,"Parkinson's","Normal")
table(trainP_data$status,pred_class) 

# A cutoff of 0.25 seems reasonable


```


## Predictions on test data

* After choosing a cutoff value, predict the class labels on the test data using our model

```{r}

prob_test <- predict(log_parkinsons_step, testP_data, type = "response")

preds_test <- ifelse(prob_test > 0.3,"Parkinson's","Normal")

table(testP_data$status,preds_test)


```

# Evaluation Metrics for classification

## Manual Computation

### Confusion Matrix

* Create a confusion matrix using the table() function

```{r}
test_data_labs <- testP_data$status

conf_matrix <- table(test_data_labs, preds_test)

print(conf_matrix)

```

### Specificity

* Calculate the Specificity

* The Proportion of correctly identified negatives by the test/model.

$${Specificity} = \frac{Number~of~True~Negatives}{Number~of~True~Negatives + Number~of~False~Positives}$$

```{r}

specificity <- conf_matrix[1, 1]/sum(conf_matrix[1, ])

print(specificity)

```


### Sensitivity

* Calculate the Sensitivity

* The Proportion of correctly identified positives by the test/model.

$${Sensitivity} = \frac{Number~of~True~Positives}{Number~of~True~Positives + Number~of~False~Negatives}$$

```{r}

sensitivity <- conf_matrix[2, 2]/sum(conf_matrix[2, ])

print(sensitivity)

```

### Accuracy

* Calculate the Accuracy

* The Proportion of correctly identified psotivies/negatives in the entire population by the test/model

$${Accuracy} = \frac{Number~of~True~Positives +Number~of~True~Negatives}{Number~Of~Subjects~in~the~Population}$$

```{r}

accuracy <- sum(diag(conf_matrix))/sum(conf_matrix)

print(accuracy)

```

## Automated Computation through Caret

* Use the caret package to compute the evaluation metrics

* Mention your reference positive class as an argument to the confusionMatrix() function

```{r}
library(e1071)
confusionMatrix(preds_test, testP_data$status, positive = "Parkinson's")


```











